---
title: "Survival Analysis"
output: html_document
date: "2025-04-29"
---

Let's do some preparation for the class by running the following chunk of codes

```{r setup}
#load packages
pacman::p_load(
  tidyverse,
  xfun,
  broom.helpers,
  survminer,
  survival,
  gtsummary
)

#load data
aset<-readRDS("C:/proj/data/als_aset2.rda") %>%
  mutate(MORT_1YR = case_when(OUTCOME_STATUS==1&DAYS_ALS1DX_TO_OUTCOME<=365~1,TRUE~0)) %>%
  mutate(SEX_F = case_when(SEX=='F'~1, TRUE~0)) %>%
  mutate(status = OUTCOME_STATUS, time = DAYS_ALS1DX_TO_OUTCOME) %>%
  filter(time > 0)
```

## Kaplan-Meier Curve

Let's first get familiar with survival object. The `Surv()` function creates a survival object that contains three parts: a numerical part that is the survival time; a `+` or `:1` symbol suggesting the subject being censored or an event. For example,

```{r}
Surv(aset$time,aset$status)[1:10]
```


```{r}
km_fit<-survfit(Surv(time,status) ~ 1, data = aset)
```

### Estimate 𝑥-year survival, or life table

```{r}
summary(km_fit,times = 365.25*c(1:5))
```


```{r}
ggsurvplot(
  fit = km_fit,
  conf.int = TRUE,
  risk.table = TRUE,
  surv.median.line = "hv", 
  break.x.by = 365*2,
  xlab = "Days", 
  ylab = "Mortality Endpoint"
)
```

### Estimate median survival time

```{r}
survfit(Surv(time, status) ~ 1, data = aset) %>% 
  tbl_survfit(
    probs = 0.5,
    label_header = "**Median survival (95% CI)**"
  )
```


### Stratified KM curves

#### Stratified KM curves by RILUZOLE_IND

```{r,fig.height=8,fig.width=12}
ggsurvplot(
    fit = survfit(Surv(time,status) ~ RILUZOLE_IND, data = aset),
    pval = TRUE, 
    conf.int = TRUE,
    risk.table = TRUE,
    linetype = "strata", 
    break.x.by = 365,
    surv.median.line = "hv",
    xlab = "Days", 
    ylab = "Mortality Endpoint"
)

```

```{r,fig.height=8,fig.width=12}
plot<-ggsurvplot(
    fit = survfit(Surv(time,status) ~ RILUZOLE_IND, data = aset),
    pval = TRUE, 
    conf.int = TRUE,
    risk.table = TRUE,
    linetype = "strata", 
    break.x.by = 365,
    xlab = "Days", 
    ylab = "Mortality Endpoint"
)

# calculate median survival
medians <- surv_median(fit = survfit(Surv(time, status) ~ SEX_F, data = aset))

# add median survival line for each group
plot$plot <- plot$plot +
  geom_vline(xintercept = medians$median, linetype = "dashed", color = "grey") 

plot
```

#### Stratified KM curves by SEX

```{r,fig.height=8,fig.width=12}
plot<-ggsurvplot(
    fit = survfit(Surv(time,status) ~ SEX_F, data = aset),
    pval = TRUE, 
    conf.int = TRUE,
    risk.table = TRUE,
    risk.table.col = "strata",
    linetype = "SEX_F", 
    break.x.by = 365,
    xlab = "Days", 
    ylab = "Mortality Endpoint"
)

# calculate median survival
medians <- surv_median(fit = survfit(Surv(time, status) ~ SEX_F, data = aset))

# add median survival line for each group
plot$plot <- plot$plot +
  geom_vline(xintercept = medians$median, linetype = "dashed", color = "grey") 

plot
```


## Cox Proportional Hazard Regression Model

```{r}
ggplot(aset,aes(x=time))+
  geom_density(aes(fill=as.factor(status)),alpha=0.5)
```


```{r}
cox_fit<-coxph(Surv(time, status) ~ RILUZOLE_IND, data = aset)
summary(cox_fit)
```


```{r}
coxph(Surv(time, status) ~ RILUZOLE_IND, data = aset) %>% 
  tbl_regression(exp = TRUE)
```


```{r}
coxph(Surv(time, status) ~ SEX_F, data = aset) %>% 
  tbl_regression(exp = TRUE)
```

```{r}
cox_fit2<-coxph(Surv(time, status) ~ AGE_AT_DX_DATE1 + RILUZOLE_IND + RILUZOLE_DUR , data = aset)
summary(cox_fit2)
```


### Model Diagnostics

```{r}
res<-cox.zph(cox_fit2)
plot(res)
```

**Add Time-Dependent Covariate (interaction with Time)**

```{r}
aset<-aset %>%
  mutate(RILUZOLE_DUR_time = RILUZOLE_DUR*log(time))

cox_fit3<-coxph(Surv(time, status) ~ AGE_AT_DX_DATE1 + RILUZOLE_IND + RILUZOLE_DUR + RILUZOLE_DUR_time, data = aset)
summary(cox_fit3)
plot(cox.zph(cox_fit3))
```



## Accelerated Failure Time Model

```{r}
aft_fit1<-survreg(Surv(time, status) ~ AGE_AT_DX_DATE1 + RILUZOLE_IND + RILUZOLE_DUR, data = aset,
              dist = "exp")
summary(aft_fit1)
```


```{r}
aft_fit2<-survreg(Surv(time, status) ~ AGE_AT_DX_DATE1 + RILUZOLE_IND + RILUZOLE_DUR, data = aset,
              dist = "weib")
summary(aft_fit2)
```


```{r}
aft_fit3<-survreg(Surv(time, status) ~ AGE_AT_DX_DATE1 + RILUZOLE_IND + RILUZOLE_DUR, data = aset,
              dist = "logl")
summary(aft_fit3)
```


```{r}
aft_fit4<-survreg(Surv(time, status) ~ AGE_AT_DX_DATE1 + RILUZOLE_IND + RILUZOLE_DUR, data = aset,
              dist = "logn")
summary(aft_fit4)
```


```{r}
AIC(aft_fit1,
    aft_fit2,
    aft_fit3,
    aft_fit4,
    cox_fit2,
    cox_fit3)
```